---
title: "TimeSeriesFeatures"
author: "Madlen Wilmes"
date: "6/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Time Series Features

Any number or vector that can be computed from a time series, is a feature of that time series. In other terms, features are numerical summaries that are useful to describe/explore time series. Some examples:
- mean, min, or max
- autocorrelation
- Guerrero estimate of Box-Cox Transformation


```{r Markdown cheatsheet, echo=FALSE}
# echo=FALSE prevents code, but not the results from appearing in the rendered file
# eval=TRUE evaluates the code and includes the results
# results='hide' hides the results but displays the code
```

### Load packages

```{r eval=TRUE, message=FALSE, warning=FALSE}
library(bigrquery)
library(ggplot2)
library(tidyr)
library(dplyr)
library(lubridate) # handle date information
library(readr) # read txt file with credentials
library(fable)
library(tsibble)
library(feasts) # stands for: Feature Extraction And Statistics for Time Series
```

## TODO --> source the ReadData or read the csv

```{r read in data from csv}
counts_day <- read.csv("crime_weather_day.csv")
head(counts_day)
# format as date
counts_day$date <- as.Date(counts_day$date)
head(counts_day)
# transform into tsibble
daily_counts <- as_tsibble(counts_day, index=date) %>% arrange(date)
```


```{r calculate avg daily counts per month}
avg_daily_counts <- daily_counts %>%
  index_by(year_month = ~ yearmonth(.)) %>% # aggregate daily averages by month
    summarise(
      avg_counts = mean(counts, na.rm = TRUE)
    )
```

```{r calculate features}
# mean of monthly aggregated avg daily counts
avg_daily_counts %>% features(avg_counts, mean)
# quantiles -> numerical equivalent of box-plot (min, lower quartile, median, upper quartile, max)
avg_daily_counts %>% features(avg_counts, quantile, prob=seq(0,1, by=0.25))
```

## Autocorrelation-based features

Reminder: autocorrelation measures the linear relationship between lagged values of a time series.

The `feat_acf()` function computes a selection of features derived from autocorrelations:

- first autocorrelation coefficient from the original data (detect non-randomness, if data with lag one is similar, they are autocorrelated);
- sum of square of the first ten autocorrelation coefficients from the original data (how much do coefficients vary around their mean --> extend of autocorrelation; how much autocorrelation there is in a series, regardless of lag)
- first autocorrelation coefficient from the differenced data;
- sum of square of the first ten autocorrelation coefficients from the differenced data;
- first autocorrelation coefficient from the twice differenced data;
- sum of square of the first ten autocorrelation coefficients from the twice differenced data;
- If seasonal data, also returns the autocorrelation coefficient at the first seasonal lag (i.e., how series changes between years, rather than between months)

```{r calculate acf features}
avg_daily_counts %>% features(avg_counts, feat_acf)
```

## STL decomposition-based features

Recap: data = trend + seasonal + remainder component

If there is a strong trend in the data, the coefficient of var(remainder) / var(trend + remainder) should be small. 
If there is a weak or no trend, the coefficient is large.
We can express the coefficient as a number between 0 and 1 to derive a measure of the "strenght of trend". Here's how to calculate it (and other features):

```{r calculate stl features}
avg_daily_counts %>% features(avg_counts, feat_stl)
```

The function `feat_stl` gives us a number of additional insight.

The strength of seasonality is defined similar to the coefficient above, but with respect to the detrended data (instead of the seasonally adjusted data) --> var(remainder) / var(seasonal + remainder). This coefficient is indicated by `seasonal_strength_year`. If it is close to zero, there's not much seasonal trend in the data. 

`seasonal_peak_year` tells us that month number 7 (July) has the peak (i.e., largest avg daily offenses). Similarily, `seasonal_trough_year` tells us that February has the lowest avg daily offenses. 

`spikiness` refers to the spikes in the remainder component (variance of the leave-one-out variances).

`stl_e_acf1` is the first autocorrelation coefficient of the remainder series (the smaller, the more random the remainder... we want to explain most by trend and seasonality and make remainder truly random)

`stl_e_acf10` is the sum of squares of the first ten autocorrelation coefficients of the remainder series; the smaller, the better, we do not want to see residual seasonality in the remainder (then we didn't do a good job with our model)

```{r calculate stl features with options}
feat_stl(avg_daily_counts$avg_counts, .period = 12, s.window = 7)
```
