---
title: "crime_weather_covid_ExploreOffenseTypes"
author: "Madlen Wilmes"
date: "6/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Load packages

```{r eval=TRUE, message=FALSE, warning=FALSE}
library(bigrquery)
library(ggplot2)
library(tidyr)
library(dplyr)
library(lubridate) # handle date information
library(readr) # read txt file with credentials
library(fable)
library(tsibble)
library(feasts) # stands for: Feature Extraction And Statistics for Time Series
```


### Retrieve Data from BigQuery
```{r get crime data}
source("crime_weather_covid_ReadData.R")
# daily_counts <- get_crime_per_day()
daily_counts_per_type <- get_crime_counts_per_type_and_day()
daily_weather <- get_weather_data()
```

```{r join and save}
# left join throws (strange) error if length of tsibble not specified
  daily_crime_weather_per_type <- left_join(daily_counts_per_type[1:nrow(daily_counts_per_type),], daily_weather, by = "date")
  write.csv(daily_crime_weather_per_type, "crime_weather_day_per_type.csv", row.names = FALSE)
```

## Read csv, if you've aleady retrieved data today
```{r read csv with crime and weather per type and day}
daily_counts_per_type <- read.csv("crime_weather_day_per_type.csv")
daily_counts_per_type$date <- as.Date(daily_counts_per_type$date)
daily_counts_per_type <- as_tsibble(daily_counts_per_type, index=date, key=c(iucr, primary_type, description)) %>% arrange(date)
```

## Counts per offense type over entire period
```{r How many offenses since 2001 per type}
daily_counts_per_type %>% group_by(iucr, primary_type,  description) %>% summarise(offense_counts = n(), .groups = 'keep')
```
Note that the same offenses have differences in the primary type and description (e.g., `CRIM SEXUAL ASSAULT` == `CRIMINAL SEXUAL ASSAULT`)
--> DATA CLEANUP!!! YAY

```{r turn into tibble}
tbl_daily_counts_type <- as_tibble(daily_counts_per_type)
```

```{r}
# unique iucr codes
tbl_daily_counts_type %>% summarise(nb_groups = n_distinct(iucr))
# unique primary_type
tbl_daily_counts_type %>% summarise(nb_groups = n_distinct(primary_type))
# unique description
tbl_daily_counts_type %>% summarise(nb_groups = n_distinct(description))
```


## Filter particular offenses
```{r}
filter(tbl_daily_counts_type, iucr == '0262')
```


## Which offense is most common over time?
```{r}
tbl_daily_counts_type %>%
  group_by(iucr, primary_type, description) %>%
  summarise(ttl_count = sum(counts, na.rm = TRUE), .groups = 'keep') %>% 
  arrange(desc(ttl_count))
```

### Which offenses occurred daily since 2001
```{r}
every_day_offenses <- daily_counts_per_type %>% has_gaps()
# how many offenses reported daily
table(every_day_offenses$.gaps)
# filter daily offenses (TRUE --> non-daily, FALSE --> daily)
filter(every_day_offenses, .gaps == FALSE)
# filter non-daily offenses (TRUE --> non-daily, FALSE --> daily)
filter(every_day_offenses, .gaps == TRUE)
```
```{r example visualization of particular offense type}
daily_counts_per_type %>% filter(iucr == '0560') %>% filter_index("2015-01-01" ~ .) %>% autoplot(counts)
```

```{r monthly averages}
daily_counts_per_type %>% 
    filter_index("2018-01-01" ~ .) %>% 
    filter(iucr == '0560') %>% 
    index_by(year_month = ~ yearmonth(.)) %>% # monthly aggregates
      summarise(avg_counts = mean(counts, na.rm = TRUE)) %>%
  as_tsibble() %>%
  autoplot(avg_counts)
```