---
title: "The influence of weather and a pandemic on crime"
output:
  html_document:
    df_print: paged
---
```{r Markdown cheatsheet, echo=FALSE}
# echo=FALSE prevents code, but not the results from appearing in the rendered file
# eval=TRUE evaluates the code and includes the results
# results='hide' hides the results but displays the code
```


### Install packages (if required)

```{r install packages, eval=FALSE} 
# set eval=TRUE if running script for the first time
# install.packages("bigrquery")
# install.packages("devtools") 
# devtools::install_github("rstats-db/bigrquery", force = TRUE)
```

### Load packages

```{r eval=TRUE, results='hide'}
library(bigrquery)
library(ggplot2)
library(tidyr)
library(dplyr)
library(lubridate) # handle date information
library(readr) # read txt file with credentials
library(tsibble) # time sensitive tibble
```

### Retrieve Google Cloud credentials
Retrieve Google Cloud project ID, which I store in an external file (not uploaded to github).
```{r set up credentials, }
# Set Google Cloud project ID here
project_name <- read_file("./GC_credentials.txt")
```

### Generate a table with counts by year and offence
 The statement needs to be written in Standard SQL. 
 More info in the [BigQuery documentation](cloud.google.com/bigquery/docs/reference/standard-sql/) 

```{r setup request}
sql <- "
SELECT 
  iucr,
  primary_type,
  description, 
  year, 
  COUNT(*) AS counts
FROM `bigquery-public-data.chicago_crime.crime` 
GROUP BY 
  iucr,
  primary_type,
  description,
  year
 ORDER BY year ASC;
"
```

```{r execute SQL call}
# Execute the query and store the result
counts_year <- query_exec(sql, project = project_name, use_legacy_sql = FALSE)
```


```{r}
# order by highest offence counts (over all years)
head(counts_year[order(counts_year$counts, decreasing = TRUE),])
```

Represent data in a tabular format that is easier to digest visually (but not great for analysis)
```{r Display in wide format}
counts_wide <- counts_year %>% spread(year, counts, fill = NA, convert = FALSE)
# display ordered by highest counts in 2020
head(counts_wide[order(counts_wide$`2020`, decreasing=TRUE),])
```

### How much data do we have?

```{r}
sql <- "SELECT COUNT(*)
FROM `bigquery-public-data.chicago_crime.crime`;"
```

```{r execute SQL call}
# Execute the query and store the result
counts <- query_exec(sql, project = project_name, use_legacy_sql = FALSE, max_pages = Inf)
print(paste("There are ", counts, " rows in the database.", sep=""))
```


## Pull in date information for more granular time series analysis

```{r setup request}
sql <- "
SELECT 
  iucr,
  primary_type,
  description, 
  date
FROM `bigquery-public-data.chicago_crime.crime` 
"
```

```{r execute SQL call}
# Execute the query and store the result
counts_day <- query_exec(sql, project = project_name, use_legacy_sql = FALSE, max_pages = Inf)
str(counts_day)
```

```{r}
head(counts_day)
counts_day_copy <- counts_day
```

```{r write out csv}
write.csv(counts_day, "counts_day.csv", row.names = FALSE)
```

```{r}
counts_day <- read.csv("counts_day.csv")
head(counts_day)
# format as date and drop the time part
counts_day$date <- as.Date(ymd_hms(counts_day$date))
head(counts_day)
```


```{r tease out date information into separate columns}
# generating the columns removes the need to create them ad hoc in every calculation below
counts_day_parts <- counts_day %>% 
  mutate(date = as.Date(date)) %>%
  mutate(year = format(date, "%y"))  %>%
  mutate(month = format(date, "%m")) %>%
  mutate(day = format(date, "%d"))
head(counts_day_parts)
```


```{r calculate the sum offense counts for each month per year}
counts_day_parts$iucr <- factor(counts_day_parts$iucr)
counts_day_parts$year <- as.integer(counts_day_parts$year)
counts_day_parts$month <- as.integer(counts_day_parts$month)
counts_day_parts$day <- as.integer(counts_day_parts$day)

# specify package of function as there are naming conflicts
per_month <- counts_day_parts %>% 
  group_by(year, month) %>% 
  dplyr::summarise(counts = dplyr::n())

head(per_month)
```

### plot monthly totals per year (across months) --> seasonal plot
```{r seasonal plot}
# generate position for labels (year 2001/2019, months Oct choosen for good position in plot)
oct_2001 <- per_month[per_month$year == 1 & per_month$month == 10,]$counts
oct_2019 <- per_month[per_month$year == 19 & per_month$month == 10,]$counts

# years <- format(seq(as.Date("2001/1/1"), by = "year", length.out = 20), "%Y")
per_month %>%
  # second month as date for axis
  mutate(month2 = as.Date(paste0("2015-", month,"-01"),"%Y-%m-%d")) %>%
  # set color value depending on year
  mutate(highlight=ifelse(year=="20", "color", "gray")) %>%
  ggplot() + 
  aes(x = month2, y = counts, group=year, color=highlight) +
  geom_line(aes(size = highlight)) +
  scale_size_manual(values=c(1.5, 0.5)) +
  scale_color_manual(values = c("#0442BF", "lightgrey"), name = "Product", labels = c('Other', '2020')) +
  labs(title = "Monthly Counts of Offenses in Chicago",
           subtitle = "Each line represents one year, reavealing the seasonality of data.",
           y = "No of Offenses",
           x = "Month") + 
  theme_minimal(base_size = 12) +
  theme(legend.position = "none") +
  # format x labels
  scale_x_date(date_labels = "%b", date_breaks = "2 months") +
  # start plot at zero
  scale_y_continuous(limits = c(0, 50000), labels = scales::number_format(accuracy = 10000, big.mark = ',')) +
  # add annotation to the plot
  geom_segment(aes(x = as.Date("2015-10-01"), y = oct_2001, xend = as.Date("2015-10-01"), yend = oct_2019), color="darkgray", size=.8, arrow = arrow(length = unit(0.1, "inches"))) +
  annotate("text", x = as.Date("2015-10-01"), y = oct_2001 + 2000, label = "Year 2001", size = 3) +
  annotate("text", x = as.Date("2015-10-01"), y = oct_2019 - 2000, label = "Year 2019", size = 3) +
  annotate("text", x = as.Date("2015-05-10"), y = 9000, label = "Year 2020", size = 3)
# ggsave("MonthlyOffenseCounts.png")
```


```{r add colum}
counts_day$count <- rep(1,nrow(counts_day))
# sort by ascending date
counts_day <- counts_day %>% arrange(date)
```

```{r aggregate counts per day}
# group by day (across offense types)
counts_day <- tibble(counts_day)

daily_counts <- test %>%
  group_by(date) %>%
  summarize(counts = sum(count))
head(daily_counts)
```

## tsibble: time-sensitive tibble
1. Index is a variable with inherent ordering from past to present.
2. Key is a set of variables that define observational units over time.
3. Each observation should be uniquely identified by index and key.
4. Each observational unit should be measured at a common interval, if regularly spaced.
```{r time-aware tibble}
# turn into time-sensitive tibble
# works like a normal tibble but tracks time (set by index)
daily_counts <- as_tsibble(daily_counts, index=date) %>% arrange(date)
```


```{r}
daily_counts %>%
ggplot() +
  aes(date, counts, color=factor(year(date)), shape=".", alpha = 0.1) +
  geom_point() +
  labs(title = "Daily Counts of Offenses in Chicago",
           subtitle = "Each dot represents one day, reavealing the seasonality of data.",
           y = "No. of Offenses",
           x = "Date") +
  # remove grid lines
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none")
```
The very granular chart reveals the downward trend of crime counts over the years, but also a strong seasonal component within each year. We also reveal daily extremes, or outliers, were the counts are a lot higher than counts on the day before or after. We can also identify the steep drop of crime rates with the onset of COVID-19 (Mar 2020).
  

# filter for specific time period
```{r filter specific days}
daily_counts %>%
    filter_index('2019-01-01' ~ '2020-05-05')%>%
    # Plotting
    ggplot() +
  aes(date, counts, color=factor(year(date))) +
  geom_point() +
  labs(title = "Daily Counts of Offenses in Chicago",
           subtitle = "Each dot represents one day, reavealing the seasonality of data.",
           y = "No. of Offenses",
           x = "Date",
           # rename the legend for color grouping
           color='Year') +
  # remove grid lines
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

# aggregate by months
```{r summarize per months}
monthly_counts <- daily_counts %>%
  index_by(year_month = ~ yearmonth(.)) %>% # monthly aggregates
  summarise(
    min_counts = min(counts, na.rm = TRUE),
    max_counts = max(counts, na.rm = TRUE),
    ttl_counts = sum(counts, na.rm = TRUE)
  )
```


```{r}
monthly_counts %>%
  mutate(ym_date = as_date(year_month)) %>%
ggplot() +
  aes(ym_date, ttl_counts, shape=".", alpha = 0.1) +
  geom_point() +
  # set titles
  labs(title = "Monthly Counts of Offenses in Chicago",
           subtitle = "Each dot represents total counts for one month. Loess smoothing emphasizes\nthe decline of offenses over the years.",
           y = "No. of Offenses",
           x = "Time") +
  # format numbers
  scale_y_continuous(labels = scales::number_format(accuracy = 10000, big.mark = ',')) +
  geom_smooth() +
  # remove grid lines
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none") + 
  # add annotation
  geom_segment(aes(x = as.Date("2020-04-01"), y = 35000, xend = as.Date("2020-04-01"), yend = 25000), color="darkgray", size=.5, arrow = arrow(length = unit(0.1, "inches"))) +
  annotate("text", x = as.Date("2020-1-01"), y = 37000, label = "COVID-19", size = 3)
```
About Loess regression:

Loess, short for Local Regression, is a non-parametric approach that fits multiple regressions in local neighborhood. It be applied using the loess() on a numerical vector to smoothen it and to predict the Y locally (i.e, within the trained values of Xs). The size of the neighborhood can be controlled using the span argument, which ranges between 0 to 1. It controls the degree of smoothing. So, the greater the value of span, more smooth is the fitted curve.

The predictor variable can just be indices from 1 to number of observations in the absence of explanatory variables. If other explanatory variables are available, they can be used as well (maximum of 4).

